---
title: "Data Visualization"
author: "Ruth Enriquez"
date: "2022-08-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(here)
library(janitor)

library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)

library(gapminder)

#library(sf)
```

### Read in the lizard data

```{r}
lizards <- read_csv(here("data_tidy", "lizards.csv"))
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point()

#you can also write this ggplot(data = lizards) + geom_point(aes(x = total_length, y = weight))
```

Create a histogram of total lengths for all lizards in the dataset.

```{r}
ggplot(data = lizards, aes (x = total_length)) +
  geom_histogram(color = "orange",
                 fill = "purple",
                 size = 0.3,
                 linetype = "dotted")
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight))+
  geom_point(
    shape = 25,
    color = "cyan4",
    fill = "yellow",
    size = 4,
    alpha = 0.4
  )
```

```{r}
ggplot(data = lizards, 
       aes(x = total_length, 
           y = weight))+
  geom_point(aes(color = common_name,
                 size = total_length),
             alpha = 0.5) +
  theme_dark()
```


```{r}
ggplot(data = lizards, 
       aes(x = total_length, 
           y = weight))+
  geom_point() +
  facet_wrap(~common_name, scales = "free")
```

##reordering

```{r}
lizard_counts <- lizards |> 
  group_by(common_name) |> 
  summarize(
    counts = n()
  )

#count - n(): will return number of rows that show up

lizard_counts2 <- lizards |> 
  count(common_name, site, tail)
```

bar plot/column graph that plots value n

```{r}
ggplot(data = lizard_counts, 
       aes(y= fct_reorder(common_name, counts), 
           x = counts)) +
  geom_col()
```

Create a scatterplots, faceted by species, for total length versus weight. The color and shape of points should differ based on the species (common name). Change the theme. Make each faceted panel have flexible axis limits (scales should be free)

```{r}
ggplot(data = lizards, aes(x = total_length,
                           y = weight)) +
  geom_point(aes(color = common_name),
             alpha = 0.6,
             size = 2) +
  theme_dark() +
  facet_wrap(~common_name, scales = "free") +
  labs(x = "Total Length (mm)", 
       y = "Weight (g)",
       title = "Jornada Basin Lizard Sizes",
       subtitle = "All lizards follow the standard lenght-weight relationships, with Western Whiptails being largest",
       caption = "Data: Jornada Basin LTER, CCBY")

#took out shapes because there was only up to 6
```







